---
title: "State level crop production"
format: html
bibliography: references.bib
---

```{r results='hide', warning=FALSE, echo=FALSE, message=FALSE}

library(tidyverse)
library(terra)
library(tidyterra)
library(readODS)
library(ggplot2)
library(ggtext)


```

# Load the States

```{r}
#| label: states
#| message: false
#| warning: false

vect_state_geo <- vect("D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/NGA Admin 2019-2025 OSGOF-Ehealth-UNCS/nga_admin1_topo.topojson")
vect_state_geo$adm1_pcode <- as.factor(vect_state_geo$adm1_pcode)
levels(vect_state_geo$adm1_pcode) # Check levels
plot(vect_state_geo, main = "Nigeria administrative level 1 boundaries (COD-AB)", levels=TRUE)

ggplot(vect_state_geo)+
  geom_spatvector(aes(fill = adm1_pcode)) +
  geom_spatvector_text(aes(label = adm1_pcode), size = 3) +
  theme_minimal()

df_state <- as.data.frame(vect_state_geo)


```



# Load the Crop Production data

## HarvestStat

```{r}
#| label: hstat_production_data
#| message: false
#| warning: false

df_hstat <- read.csv("D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/HarvestStat/adm_crop_production_NG.csv")
df_hstat$admin_1 <- as.factor(df_hstat$admin_1)
levels(df_hstat$admin_1) # Check levels


```

The HarvestStat dataset does not contain codes for the states - only names.
Therefore we need to join the two data frames using the name of the state - and checking that all states have data.

```{r}
#| label: hstat_name_join
#| message: false
#| warning: false

df_state_production <- dplyr::left_join(df_state, df_hstat, by = c("adm1_name" = "admin_1"))
str(df_state_production)

```


### Rice

```{r}
#| label: hstat_map_test_rice
#| message: false
#| warning: false

vect_rice_2023 <- dplyr::left_join(vect_state_geo, dplyr::filter(df_hstat, product == "Rice", harvest_year == "2023", indicator             == "area"), by = c("adm1_name" = "admin_1"))

df_rice_2023 <- as.data.frame(vect_rice_2023)
df_rice_2023


ggplot(vect_rice_2023) +
  geom_spatvector(aes(fill = value)) +
  geom_spatvector_text(aes(label = adm1_pcode), size = 2) +
  scale_fill_stepsn(
      n.breaks = 10,
      limits = c(0, 400000),
      colours = hcl.colors(palette = "Grays", 10, rev = T),
      guide = "legend",  na.value = "pink", labels = scales::comma
    ) +
  labs(
    title = "Rice Production - 2023",
    fill = "Cropped<br>Area (ha)") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

### Maize

```{r}
#| label: hstat_map_test_maize
#| message: false
#| warning: false

vect_maize_2023 <- dplyr::left_join(vect_state_geo, dplyr::filter(df_hstat, product == "Maize", harvest_year == "2023", indicator             == "area"), by = c("adm1_name" = "admin_1"))

df_maize_2023 <- as.data.frame(vect_maize_2023)
df_maize_2023


ggplot(vect_maize_2023) +
  geom_spatvector(aes(fill = value)) +
  geom_spatvector_text(aes(label = adm1_pcode), size = 2) +
  scale_fill_stepsn(
      n.breaks = 10,
      limits = c(0, 400000),
      colours = hcl.colors(palette = "Grays", 10, rev = T),
      guide = "legend",  na.value = "pink", labels = scales::comma
    ) +
  labs(
    title = "Maize Production - 2023",
    fill = "Cropped<br>Area (ha)") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

### Cassava

```{r}
#| label: hstat_map_test_cassava
#| message: false
#| warning: false

vect_cassava_2023 <- dplyr::left_join(vect_state_geo, dplyr::filter(df_hstat, product == "Cassava", harvest_year == "2023", indicator == "area"), by = c("adm1_name" = "admin_1"))

df_cassava_2023 <- as.data.frame(vect_cassava_2023)
df_cassava_2023


ggplot(vect_cassava_2023) +
  geom_spatvector(aes(fill = value)) +
  geom_spatvector_text(aes(label = adm1_pcode), size = 2) +
  scale_fill_stepsn(
      n.breaks = 10,
      limits = c(0, 1000000),
      colours = hcl.colors(palette = "Grays", 10, rev = T),
      guide = "legend",  na.value = "pink", labels = scales::comma
    ) +
  labs(
    title = "Cassava Production - 2023",
    fill = "Cropped<br>Area (ha)") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

### Sweet Potato

```{r}
#| label: hstat_map_test_spotato
#| message: false
#| warning: false

vect_spotato_2006 <- dplyr::left_join(vect_state_geo, dplyr::filter(df_hstat, product == "Sweet Potatoes", harvest_year == "2006", indicator == "area"), by = c("adm1_name" = "admin_1"))

df_spotato_2006 <- as.data.frame(vect_spotato_2006)
df_spotato_2006


ggplot(vect_spotato_2006) +
  geom_spatvector(aes(fill = value)) +
  geom_spatvector_text(aes(label = adm1_pcode), size = 2) +
  scale_fill_stepsn(
      n.breaks = 10,
      limits = c(0, 30000),
      colours = hcl.colors(palette = "Grays", 10, rev = T),
      guide = "legend",  na.value = "pink", labels = scales::comma
    ) +
  labs(
    title = "Sweet Potato Production - 2006",
    fill = "Cropped<br>Area (ha)") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

### Cowpea

```{r}
#| label: hstat_map_test_cowpea
#| message: false
#| warning: false

vect_cowpea_2023 <- dplyr::left_join(vect_state_geo, dplyr::filter(df_hstat, product == "Cowpea", harvest_year == "2023", indicator == "area"), by = c("adm1_name" = "admin_1"))

df_cowpea_2023 <- as.data.frame(vect_cowpea_2023)
df_cowpea_2023


ggplot(vect_cowpea_2023) +
  geom_spatvector(aes(fill = value)) +
  geom_spatvector_text(aes(label = adm1_pcode), size = 2) +
  scale_fill_stepsn(
      n.breaks = 10,
      limits = c(0, 500000),
      colours = hcl.colors(palette = "Grays", 10, rev = T),
      guide = "legend",  na.value = "pink", labels = scales::comma
    ) +
  labs(
    title = "Cowpea Production - 2023",
    fill = "Cropped<br>Area (ha)") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```



### Sorghum

```{r}
#| label: hstat_map_test_sorghum
#| message: false
#| warning: false

vect_sorghum_2023 <- dplyr::left_join(vect_state_geo, dplyr::filter(df_hstat, product == "Sorghum", harvest_year == "2023", indicator == "area"), by = c("adm1_name" = "admin_1"))

df_sorghum_2023 <- as.data.frame(vect_sorghum_2023)
df_sorghum_2023


ggplot(vect_sorghum_2023) +
  geom_spatvector(aes(fill = value)) +
  geom_spatvector_text(aes(label = adm1_pcode), size = 2) +
  scale_fill_stepsn(
      n.breaks = 10,
      limits = c(0, 500000),
      colours = hcl.colors(palette = "Grays", 10, rev = T),
      guide = "legend",  na.value = "pink", labels = scales::comma
    ) +
  labs(
    title = "Sorghum Production - 2023",
    fill = "Cropped<br>Area (ha)") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

### Banana

```{r}
#| label: hstat_map_test_banana
#| message: false
#| warning: false

vect_banana_2021 <- dplyr::left_join(vect_state_geo, dplyr::filter(df_hstat, product == "Banana", harvest_year == "2021", indicator == "area"), by = c("adm1_name" = "admin_1"))

df_banana_2021 <- as.data.frame(vect_banana_2021)
df_banana_2021


ggplot(vect_banana_2021) +
  geom_spatvector(aes(fill = value)) +
  geom_spatvector_text(aes(label = adm1_pcode), size = 2) +
  scale_fill_stepsn(
      n.breaks = 10,
      limits = c(0, 200000),
      colours = hcl.colors(palette = "Grays", 10, rev = T),
      guide = "legend",  na.value = "pink", labels = scales::comma
    ) +
  labs(
    title = "Banana Production - 2021",
    fill = "Cropped<br>Area (ha)") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```


## NAERLS/FMAFS

#### Rice

```{r}
#| label: naerls_rice_production
#| message: false
#| warning: false

df_naerls_rice <- read.csv("D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/NAERLS Crop Stats 2024/Rice.csv")
df_naerls_rice$State <- as.factor(df_naerls_rice$State)
levels(df_naerls_rice$State) # Check levels

vect_rice_2024 <- dplyr::left_join(vect_state_geo, df_naerls_rice, by = c("adm1_name" = "State"))

df_rice_2024 <- as.data.frame(vect_rice_2024)
df_rice_2024


ggplot(vect_rice_2024) +
  geom_spatvector(aes(fill = Area_Ha_2024)) +
  geom_spatvector_text(aes(label = adm1_pcode), size = 2) +
  scale_fill_stepsn(
      n.breaks = 10,
      limits = c(0, 500000),
      colours = hcl.colors(palette = "Grays", 10, rev = T),
      guide = "legend",  na.value = "pink", labels = scales::comma
    ) +
  labs(
    title = "Rice Production - 2024",
    fill = "Cropped<br>Area (ha)") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

#### Maize

```{r}
#| label: naerls_maize_production
#| message: false
#| warning: false

df_naerls_maize <- read.csv("D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/NAERLS Crop Stats 2024/Maize.csv")
df_naerls_maize$State <- as.factor(df_naerls_maize$State)
levels(df_naerls_maize$State) # Check levels

vect_maize_2024 <- dplyr::left_join(vect_state_geo, df_naerls_maize, by = c("adm1_name" = "State"))

df_maize_2024 <- as.data.frame(vect_maize_2024)
df_maize_2024


ggplot(vect_maize_2024) +
  geom_spatvector(aes(fill = Area_Ha_2024)) +
  geom_spatvector_text(aes(label = adm1_pcode), size = 2) +
  scale_fill_stepsn(
      n.breaks = 10,
      limits = c(0, 500000),
      colours = hcl.colors(palette = "Grays", 10, rev = T),
      guide = "legend",  na.value = "pink", labels = scales::comma
    ) +
  labs(
    title = "Maize Production - 2024",
    fill = "Cropped<br>Area (ha)") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

#### Cassava

```{r}
#| label: naerls_cassava_production
#| message: false
#| warning: false

df_naerls_cassava <- read.csv("D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/NAERLS Crop Stats 2024/Cassava.csv")
df_naerls_cassava$State <- as.factor(df_naerls_cassava$State)
levels(df_naerls_cassava$State) # Check levels

vect_cassava_2024 <- dplyr::left_join(vect_state_geo, df_naerls_cassava, by = c("adm1_name" = "State"))

df_cassava_2024 <- as.data.frame(vect_cassava_2024)
df_cassava_2024


ggplot(vect_cassava_2024) +
  geom_spatvector(aes(fill = Area_Ha_2024)) +
  geom_spatvector_text(aes(label = adm1_pcode), size = 2) +
  scale_fill_stepsn(
      n.breaks = 10,
      limits = c(0, 1000000),
      colours = hcl.colors(palette = "Grays", 10, rev = T),
      guide = "legend",  na.value = "pink", labels = scales::comma
    ) +
  labs(
    title = "Cassava Production - 2024",
    fill = "Cropped<br>Area (ha)") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

#### Cowpea

```{r}
#| label: naerls_cowpea_production
#| message: false
#| warning: false

df_naerls_cowpea <- read.csv("D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/NAERLS Crop Stats 2024/Cowpea.csv")
df_naerls_cowpea$State <- as.factor(df_naerls_cowpea$State)
levels(df_naerls_cowpea$State) # Check levels

vect_cowpea_2024 <- dplyr::left_join(vect_state_geo, df_naerls_cowpea, by = c("adm1_name" = "State"))

df_cowpea_2024 <- as.data.frame(vect_cowpea_2024)
df_cowpea_2024


ggplot(vect_cowpea_2024) +
  geom_spatvector(aes(fill = Area_Ha_2024)) +
  geom_spatvector_text(aes(label = adm1_pcode), size = 2) +
  scale_fill_stepsn(
      n.breaks = 10,
      limits = c(0, 1000000),
      colours = hcl.colors(palette = "Grays", 10, rev = T),
      guide = "legend",  na.value = "pink", labels = scales::comma
    ) +
  labs(
    title = "Cowpea Production - 2024",
    fill = "Cropped<br>Area (ha)") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

#### Sorghum

```{r}
#| label: naerls_sorghum_production
#| message: false
#| warning: false

df_naerls_sorghum <- read.csv("D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/NAERLS Crop Stats 2024/Sorghum.csv")
df_naerls_sorghum$State <- as.factor(df_naerls_sorghum$State)
levels(df_naerls_sorghum$State) # Check levels

vect_sorghum_2024 <- dplyr::left_join(vect_state_geo, df_naerls_sorghum, by = c("adm1_name" = "State"))

df_sorghum_2024 <- as.data.frame(vect_sorghum_2024)
df_sorghum_2024


ggplot(vect_sorghum_2024) +
  geom_spatvector(aes(fill = Area_Ha_2024)) +
  geom_spatvector_text(aes(label = adm1_pcode), size = 2) +
  scale_fill_stepsn(
      n.breaks = 10,
      limits = c(0, 500000),
      colours = hcl.colors(palette = "Grays", 10, rev = T),
      guide = "legend",  na.value = "pink", labels = scales::comma
    ) +
  labs(
    title = "Sorghum Production - 2024",
    fill = "Cropped<br>Area (ha)") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

#### Banana/Plantain

```{r}
#| label: naerls_plantain_banana_production
#| message: false
#| warning: false

df_naerls_plantain_banana <- read.csv("D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/NAERLS Crop Stats 2024/Plantain_Banana.csv")
df_naerls_plantain_banana$State <- as.factor(df_naerls_plantain_banana$State)
levels(df_naerls_plantain_banana$State) # Check levels

vect_plantain_banana_2024 <- dplyr::left_join(vect_state_geo, df_naerls_plantain_banana, by = c("adm1_name" = "State"))

df_plantain_banana_2024 <- as.data.frame(vect_plantain_banana_2024)
df_plantain_banana_2024


ggplot(vect_plantain_banana_2024) +
  geom_spatvector(aes(fill = Area_Ha_2024)) +
  geom_spatvector_text(aes(label = adm1_pcode), size = 2) +
  scale_fill_stepsn(
      n.breaks = 10,
      limits = c(0, 200000),
      colours = hcl.colors(palette = "Grays", 10, rev = T),
      guide = "legend",  na.value = "pink", labels = scales::comma
    ) +
  labs(
    title = "Plantain/Banana Production - 2024",
    fill = "Cropped<br>Area (ha)") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```